### 2024-04-10 diagrams/data-model.tex

\begin{tikzpicture}

 %% sample_id
  \node at (2.0,5.9,-.5) {\ttfamily \scriptsize sample\textunderscore id};
  % bounds
  \draw[densely dashed] (0,5.5,-.5) -- (3.9,5.5,-.5);
  \draw[densely dashed] (0,5.7,-.5) -- (3.9,5.7,-.5);
 \tikzcuboidset{front face/.style={fill=yellow!20},right face/.style={fill=green!10},top face/.style={fill=yellow!20}}
  \pic at (0,5.5,-.5) {cuboid=.9--.2--0};
  \pic at (1,5.5,-.5) {cuboid=.9--.2--0};
  \node at (2.5,5.6,-.5) {\ttfamily \dots};
  \pic at (3,5.5,-.5) {cuboid=.9--.2--0};

 %% variant_contig
  \node[rotate=90] at (-1.5,2.5,-.3) {\ttfamily \scriptsize variant\textunderscore contig};
  % bounds
  \draw[densely dashed] (-1.4,0,-.5) -- (-1.4,4.9,-.5);
  \draw[densely dashed] (-1.2,0,-.5) -- (-1.2,4.9,-.5);
 \tikzcuboidset{front face/.style={fill=green!10},right face/.style={fill=green!10},top face/.style={fill=yellow!20}}
  \pic at (-1.4,0,-.5) {cuboid=.2--.9--0};
  \node at (-1.3,1.5,-.5) {\ttfamily \vdots};
  \pic at (-1.4,2,-.5) {cuboid=.2--.9--0};
  \pic at (-1.4,3,-.5) {cuboid=.2--.9--0};
  \pic at (-1.4,4,-.5) {cuboid=.2--.9--0};

 %% variant_position
  \node[rotate=90] at (-.8,2.5,-.3) {\ttfamily \scriptsize variant\textunderscore position};
  % bounds
  \draw[densely dashed] (-.7,0,-.5) -- (-.7,4.9,-.5);
  \draw[densely dashed] (-.5,0,-.5) -- (-.5,4.9,-.5);
 \tikzcuboidset{front face/.style={fill=green!10},right face/.style={fill=green!10},top face/.style={fill=yellow!20}}
  \pic at (-.7,0,-.5) {cuboid=.2--.9--0};
  \node at (-.6,1.5,-.5) {\ttfamily \vdots};
  \pic at (-.7,2,-.5) {cuboid=.2--.9--0};
  \pic at (-.7,3,-.5) {cuboid=.2--.9--0};
  \pic at (-.7,4,-.5) {cuboid=.2--.9--0};

 %% call_genotype
  \node at (2.0,5.3) {\ttfamily \scriptsize call\textunderscore genotype};
  % back box
  \draw[densely dashed] (0,0,-0.5) -- (3.9,0,-0.5);
  \draw[densely dashed] (0,0,-0.5) -- (0,4.9,-0.5);
  \draw[densely dashed] (0,4.9,-0.5) -- (3.9,4.9,-0.5);
  \draw[densely dashed] (3.9,0,-0.5) -- (3.9,4.9,-0.5);
 \tikzcuboidset{front face/.style={fill=blue!10},right face/.style={fill=green!10},top face/.style={fill=yellow!20}}
  \pic at (0,0) {cuboid=.9--.9--.5};
  \node at (0.45,0.5) {\ttfamily \tiny 100/0/0};
  \node at (0.5,1.6) {\ttfamily \vdots};
  \pic at (0,2) {cuboid=.9--.9--.5};
  \node at (0.45,2.5) {\ttfamily \tiny 2/0/0};
  \pic at (0,3) {cuboid=.9--.9--.5};
  \node at (0.45,3.5) {\ttfamily \tiny 1/0/0};
  \pic at (0,4) {cuboid=.9--.9--.5};
  \node at (0.45,4.5) {\ttfamily \tiny 0/0/0};
  \pic at (1,0) {cuboid=.9--.9--.5};
  \node at (1.45,0.5) {\ttfamily \tiny 100/1/0};
  \node at (1.5,1.6) {\ttfamily \vdots};
  \pic at (1,2) {cuboid=.9--.9--.5};
  \node at (1.45,2.5) {\ttfamily \tiny 2/1/0};
  \pic at (1,3) {cuboid=.9--.9--.5};
  \node at (1.45,3.5) {\ttfamily \tiny 1/1/0};
  \pic at (1,4) {cuboid=.9--.9--.5};
  \node at (1.45,4.5) {\ttfamily \tiny 0/1/0};
  \node at (2.5,0.5) {\ttfamily \dots};
  \node at (2.5,1.55) {\ttfamily $\ddots$};
  \node at (2.5,2.5) {\ttfamily \dots};
  \node at (2.5,3.5) {\ttfamily \dots};
  \node at (2.5,4.5) {\ttfamily \dots};
  \pic at (3,0) {cuboid=.9--.9--.5};
  \node at (3.45,0.5) {\ttfamily \tiny 100/10/0};
  \node at (3.5,1.6) {\ttfamily \vdots};
  \pic at (3,2) {cuboid=.9--.9--.5};
  \node at (3.45,2.5) {\ttfamily \tiny 2/10/0};
  \pic[ultra thick, red] at (3,3) {cuboid=.9--.9--.5};
  \node at (3.45,3.5) {\ttfamily \tiny 1/10/0};
  \pic at (3,4) {cuboid=.9--.9--.5};
  \node at (3.45,4.5) {\ttfamily \tiny 0/10/0};
  % front box
  \draw[densely dashed] (0,0) -- (3.9,0);
  \draw[densely dashed] (0,0) -- (0,4.9);
  \draw[densely dashed] (0,4.9) -- (3.9,4.9);
  \draw[densely dashed] (3.9,0) -- (3.9,4.9);

 %% dimensions
  \draw[stealth-stealth] (0.1,-0.2) -- (3.8,-0.2) node[midway,below] {\scriptsize samples};
  \draw[stealth-stealth] (4,-0.2, 0) -- (4,-0.2,-.5) node[midway,below right] {\scriptsize ploidy};
  \draw[stealth-stealth] (4.1,.1,-.5) -- (4.1,4.8,-.5) node[midway,below,rotate=90] {\scriptsize variants};

%% compresion
 \node (numcodecs) [rectangle, draw, densely dotted, thick, rounded corners, minimum width=2.4cm, minimum height=2.6cm, anchor= south west]
    at (4.8,2.2) {};
 \node[anchor=north,below] at (numcodecs.north) {\scriptsize Filter \& compress};
 %\node[anchor=south,below] at (numcodecs.south) {numcodecs};
 \node (blosk) [rectangle, ultra thick, red, draw, rounded corners, minimum width=1.8cm, minimum height=1.6cm, anchor=center,align=center,text=black]
    at (numcodecs) {\scriptsize Blosc\\\scriptsize ZSTD\\\scriptsize Autoshuffle};
 \draw[ultra thick, red,-stealth] (3.9,3.5,-.25) -- (5,3.5,-.25);

%% storage
 \node (location) [rectangle, draw, densely dotted, thick, rounded corners, minimum width=2.4cm, minimum height=4.5cm, anchor= south west]
    at (7.4,1.35) {};
 \node[anchor=south,below] at (location.south) {Storage};
 \node[anchor=center] at (location) {
        \begin{forest}
          for tree={
            font=\tiny ,
            grow'=0,
            child anchor=west,
            parent anchor=south,
            anchor=west,
            calign=first,
            edge path={
              \noexpand\path [draw, \forestoption{edge}]
              (!u.south west) +(7.5pt,0) |- node[fill,inner sep=1.25pt] {} (.child anchor)\forestoption{edge label};
            },
            before typesetting nodes={
              if n=1
                {insert before={[,phantom]}}
                {}
            },
            fit=band,
            before computing xy={l=15pt},
          }
        [{\scriptsize Cloud, file or blob}
          [\ttfamily call\textunderscore genotype
            [\ttfamily 0/0/0]
            [\ttfamily ...]
            [\ttfamily 1/10/0, draw={red,ultra thick, rounded corners}]
            [\ttfamily ...]
            [\ttfamily 100/10/0]
          ]
          [\ttfamily sample\textunderscore id]
          [\ttfamily ...]
        ]
        \end{forest}
 };
 \draw[ultra thick, red, -stealth,rounded corners=5pt] (6.8,3.5,-.25) -- (8,3.5,-.25);

\end{tikzpicture}