### 2024-06-07 paper.tex
Add point about annotations
Closes #141

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please note that whilst this template provides a
% preview of the typeset manuscript for submission, it
% will not necessarily be the final publication layout.
%
% letterpaper/a4paper: US/UK paper size toggle
% num-refs/alpha-refs: numeric/author-year citation and bibliography toggle

%\documentclass[letterpaper]{oup-contemporary}
\documentclass[a4paper,num-refs]{oup-contemporary}

%%% Journal toggle; only specific options recognised.
%%% (Only "gigascience" and "general" are implemented now. Support for other journals is planned.)
\journal{gigascience}

\usepackage{graphicx}
\usepackage{siunitx}

% Additional packages for data model figure
\usepackage{forest}
\usepackage{tikz}
\usetikzlibrary{quotes,arrows.meta,3d}

% Local definitions to systematise tool naming
\newcommand{\sgapi}[1]{\texttt{#1}}
\newcommand{\toolname}[1]{\texttt{#1}}
\newcommand{\sgkit}{\texttt{sgkit}}

%%% Flushend: You can add this package to automatically balance the final page, but if things go awry (e.g. section contents appearing out-of-order or entire blocks or paragraphs are coloured), remove it!
% \usepackage{flushend}

% Macros etc for data model figure
\input{diagrams/data-model-preamble.tex}

% See https://github.com/sgkit-dev/vcf-zarr-publication/issues/87
% for discussion
\title{Analysis-ready VCF at Biobank scale using Zarr}

%%% Use the \authfn to add symbols for additional footnotes, if any. 
% 1 is reserved for correspondence emails; then continuing with 2 etc for contributions.
% First author
\author[1,\authfn{1}]{Eric Czech} % https://orcid.org/0000-0002-4254-4255
\author[2,3\authfn{1}]{Timothy R. Millar} % https://orcid.org/0000-0002-5142-8811
\author[4,\authfn{1}]{Tom White}

% Middle
\author[5]{Ben Jeffery} % https://orcid.org/0000-0002-1982-6801
\author[6]{Alistair Miles} % https://orcid.org/0000-0001-9018-4680
\author[7]{Sam Tallman} % https://orcid.org/0000-0001-7183-6276
\author[1]{Rafal Wojdyla} % https://orcid.org/0009-0005-0735-7090
\author[8]{Shadi Zabad} % https://orcid.org/0000-0002-8003-9284

% Senior 
\author[1,\authfn{2}]{Jeff Hammerbacher} % https://orcid.org/0000-0001-6596-8563
\author[5,\authfn{2},\authfn{3}]{Jerome Kelleher} % https://orcid.org/0000-0002-7894-5253

\affil[1]{Related Sciences}
\affil[2]{The New Zealand Institute for Plant \& Food Research Ltd, Lincoln,
New Zealand}
\affil[3]{Department of Biochemistry, School of Biomedical Sciences, University of Otago, Dunedin, New Zealand}
\affil[4]{Tom White Consulting Ltd.}
\affil[5]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, 
University of Oxford, UK}
\affil[6]{Wellcome Sanger Institute}
\affil[7]{Genomics England}
\affil[8]{School of Computer Science, McGill University, Montreal, QC, Canada}

%%% Author Notes
\authnote{\authfn{1}Joint first author.}
\authnote{\authfn{2}Joint senior author.}
\authnote{\authfn{3}jerome.kelleher@bdi.ox.ac.uk}

%%% Paper category
\papercat{Paper}

%%% "Short" author for running page header
\runningauthor{Czech et al.}

%%% Should only be set by an editor
\jvolume{00}
\jnumber{0}
\jyear{2024}

\begin{document}

\begin{frontmatter}
\maketitle

% The Abstract (250 words maximum) should be structured to
% include the following details:
% \textbf{Background}, the context and purpose of
% the study;
% \textbf{Results}, the main findings;
% \textbf{Conclusions}, brief
% summary and potential implications. Please minimize the use of abbreviations
% and do not cite references in the abstract.
% The Abstract (250 words maximum) should be structured to
% include the following details:
% \textbf{Background}, the context and purpose of % the study;
% \textbf{Results}, the main findings;
% \textbf{Conclusions}, brief summary and potential implications.

%% NOTE: this is much too long currently, but keeping for now so we
% can see which stuff migrates to the intro
\begin{abstract}
\textbf{Background:}
Variant Call Format (VCF) is the standard file format for interchanging
genetic variation data and associated quality control metrics.
% It provides a well-defined data model and is central to a large ecosystem
% of interoperating tools.
The usual row-wise encoding of the VCF data model (either as text
or packed binary) emphasises efficient retrieval of all data for a given
variant, but accessing data on a field or sample basis is inefficient.
Biobank scale datasets currently available 
consist of hundreds of thousands of whole genomes 
and hundreds of terabytes of compressed VCF.
Row-wise data storage is fundamentally unsuitable
and a more scalable approach is needed.

\textbf{Results:}
We present the VCF Zarr specification, an encoding of the 
VCF data model using Zarr which makes retrieving subsets of the 
data much more efficient. Zarr is a cloud-native format for storing 
multi-dimensional data, widely used in scientific computing.
We show how this format is far more efficient than
standard VCF based approaches,
and competitive with specialised methods for 
storing genotype data in terms of compression ratios
and calculation performance.
We demonstrate the VCF Zarr format (and the vcf2zarr conversion utility) 
on a subset of the Genomics England aggV2 dataset comprising
78,195 samples and 59,880,903 variants,
with a 5X reduction in storage and greater than 300X reduction in CPU usage
in some representative benchmarks.

\textbf{Conclusions:}
Large row-encoded VCF files are a major bottleneck for current research, and 
storing and processing these files incurs a substantial cost.
The VCF Zarr specification, building on widely-used, open-source technologies
has the potential to greatly reduce these costs, 
and may enable a diverse ecosystem of next-generation tools for analysing 
genetic variation data directly from cloud-based object stores.
\end{abstract}

\begin{keywords}
Variant Call Format; Zarr; Analysis ready data.
\end{keywords}
\end{frontmatter}

%%% Key points will be printed at top of second page
\begin{keypoints*}
\begin{itemize}
\item VCF is widely supported, and the underlying data model entrenched 
in bioinformatics pipelines.
\item The standard row-wise encoding as text (or binary) is inherently
inefficient for large-scale data processing.
\item The Zarr format provides an efficient solution, by encoding fields
in the VCF separately in chunk-compressed binary format.
\end{itemize}
\end{keypoints*}

\section{Background}
Variant Call Format (VCF) is the standard format for interchanging genetic
variation data, encoding information about 
DNA sequence polymorphisms among a set of samples with associated 
quality control metrics and metadata~\citep{danecek2011variant}. 
Originally defined specifically as a text file,
it has been refined and standardised~\citep{rehm2021ga4gh} and the 
underlying data-model is now deeply embedded in bioinformatics practice.
Dataset sizes have grown explosively since the introduction of 
VCF as part of 1000 Genomes project~\citep{10002015global},
with Biobank scale initiatives such as 
Genomics England~\cite{turnbull2018100},
UK Biobank~\citep{bycroft2018genome,backman2021exome,halldorsson2022sequences,uk2023whole},
and the All of Us research program~\citep{all2024genomic}
collecting genome sequence data for hundreds of thousands of humans.
Large genetic variation datasets are also being generated for other organisms
and a variety of purposes including 
agriculture~\citep{ros2020accuracy,wang2023rice}, 
conservation~\citep{shaffer2022landscape}
and infectious disease surveillance~\citep{hamid2023pf7}.
VCF's simple text-based design and widespread
support~\cite{garrison2022spectrum} makes it an 
excellent archival format, but it is an inefficient basis for analysis.
Methods that require efficient access to genotype data
either require conversion to the
PLINK~\cite{purcell2007plink,chang2015second} 
or BGEN~\citep{band2018bgen} 
formats~\citep[e.g.][]{yang2011gcta,mbatchou2021computationally,loh2015efficient}
or use bespoke binary formats 
that support the required access patterns~\citep[e.g.][]{
% Uses custom "bref3" format,
% https://faculty.washington.edu/browning/beagle/bref3.24May18.pdf
browning2018one, 
% .samples Zarr format
kelleher2019inferring,
% Has a "xcftools" package, but it still looks pretty experimental
hofmeister2023accurate}.
While PLINK and BGEN formats are more efficient to access than VCF, neither
can accommodate the full flexibility of the VCF data model and conversion
is lossy.
PLINK's approach of storing the genotype matrix in uncompressed
packed-binary format provides efficient access to genotype 
data, but file sizes are substantially larger than 
the equivalent compressed VCF (see Fig~\ref{fig-data-storage}).
For example, at two bits 
per diploid genotype, the full genotype matrix for the GraphTyper SNP dataset 
in the 500K UKB WGS data~\citep{uk2023whole} is 116 TiB.
% 1,037,556,156 SNPs x  490,640 samples
% humanize.naturalsize(1_037_556_156 * 490_640 / 4, binary=True)
% '115.7 TiB'

Processing of Biobank scale datasets can be split into a 
few broad categories. The most basic analysis 
is quality control (QC). Variant QC is an 
involved and multi-faceted 
task~\cite{marees2018tutorial,panoutsopoulou2018quality,chen2024genomic}, 
often requiring interactive, exploratory analysis
and incurring substantial computation over multiple QC fields.
Genotype calls are sometimes refined via statistical methods,
for example by phasing~\citep{
browning2021fast,browning2023statistical,hofmeister2023accurate,williams2024phasing},
and imputation~\citep{browning2018one,rubinacci2020genotype,barton2021whole,
rubinacci2023imputation}
creating additional dataset copies.
A common task to perform  is
a genome wide association study (GWAS)~\cite{uffelmann2021genome}. 
The majority of tools for 
performing GWAS and related analyses require
data to be in PLINK or BGEN formats~\cite[e.g][]{chang2015second,
loh2015efficient,
abraham2017flashpca2,
mbatchou2021computationally},
and so data must be ``hard-called'' according to some QC criteria
and exported to additional copies.
Finally, variation datasets are often queried in exploratory
analyses, to find regions or samples of interest for a particular 
study~\cite[e.g.][]{chen2024novo}.

VCF cannot support any of these workflows efficiently at the Biobank scale.
The most intrinsically limiting aspect of VCF's design 
is its row-wise layout of data, which means that (for example)
information for a particular sample or field cannot be obtained without
retrieving the entire dataset.
The file-oriented paradigm is also unsuited to the realities 
of modern datasets, which are too large to download and 
often required to stay in-situ by data-access agreements.
Large files are currently stored in cloud environments, where the 
file systems that are required by classical file-oriented tools
are expensively emulated on the basic building blocks
of object storage. 
These multiple layers of inefficiencies around processing
VCF data at scale in the cloud mean that it is 
time-consuming and expensive, and these vast datasets are 
not utilised to their full potential.

To achieve this full potential we 
need a new generation of tools that operate directly
on a primary data representation that supports 
efficient access across a range of applications,
with native support for cloud object storage.
Such a representation can be termed ``analysis-ready''
and ``cloud-native''~\citep{abernathey2021cloud}.
For the representation to be FAIR~\citep{wilkinson2016fair},
it must also be \emph{accessible}, using protocols that are 
``open, free, and universally implementable''.
There is currently no efficient, FAIR representation of genetic variation
data suitable for cloud deployments.
Hail~\cite{ganna2016ultra,hail2024} has become the dominant platform
for quality control of large-scale variation datasets, 
and has been instrumental in projects such as 
gnomadAD~\cite{karczewski2020mutational,chen2024genomic}.
Glow~\cite{glow2024} has recently emerged as a similar approach
to performing distributed analytics on genetic variation data.
While Hail and Glow are built on open components
from the Hadoop distributed computing ecosystem~\citep{white2012hadoop},
they are both complex projects consisting of 
tightly coupled components that are not intended for external reuse.
Similarly, commercial solutions that have emerged to facilitate
the analysis of large-scale genetic variation data are either
based on proprietary~\cite{basespace2024,graf2024,googlelifesciences2024,
awshealthomics2024,microsoftgenomics2024}
or single-vendor technologies~\cite[e.g.][]{tiledb2024,genomicsdb2024}.
The next generation of VCF analysis methods requires
an open, free and transparent data representation 
with multiple independent implementations.

In this article, we decouple the VCF data model from its row-oriented
file definition, and show how the data can be 
compactly stored and efficiently analysed in a cloud-native, FAIR manner.
We do this by translating VCF data into Zarr format,
a method of storing large-scale multidimensional data as a regular
grid of compressed chunks. 
Zarr's elegant simplicity and first-class support for 
cloud object stores have led to 
it gaining substantial traction
across the sciences, and it is now used in multiple petabyte-scale
datasets in cloud deployments (see Methods for details).
We present the VCF Zarr specification that formalises this 
mapping, and the \texttt{vcf2zarr} 
utility to reliably convert large-scale VCFs to Zarr.
We show that VCF Zarr is much more compact than 
VCF and is competitive with state-of-the-art
file-based VCF compression tools. 
Moreover, we show that Zarr's storage of data in an analysis-ready 
format greatly facilitates computation,
with various benchmarks being substantially faster than
\texttt{bcftools} based pipelines, and again competitive
with state-of-the-art file-oriented methods. Finally, we show the 
utility of VCF Zarr on the Genomics England aggV2 dataset,
demonstrating that common \texttt{bcftools} queries can be performed orders
of magnitude more quickly using simple Python scripts.

\section{Results}

\subsection{Storing genetic variation data}
Although VCF is the standard format for exchanging genetic variation
data, its limitations both in terms of compression 
and query/compute performance are well 
known~\citep[e.g.][]{kelleher2013processing,layer2016efficient,li2016bgt},
and many methods 
have been suggested to improve on these properties.
Most approaches balance compression with
performance on particular types of queries, 
typically using a command line interface (CLI)
and outputting VCF text~\citep{
layer2016efficient, %GQT
li2016bgt, % BGT
tatwawadi2016gtrac, % GTRAC
danek2018gtc, % GTC
lin2020sparse, % SpVCF
lan2020genozip,lan2021genozip, %genozip
lefaive2021sparse, % SAVVY
wertenbroek2022xsi,% XSI
zhang2023gbc}. %GBC
Several specialised algorithms for compressing 
the genotype matrix (i.e., just the genotype calls without additional
VCF information) have been proposed
\citep{qiao2012handling, % SpeedGene
deorowicz2013genome, %TGC
sambo2014compression, % snpack
deorowicz2019gtshark, %GTShark
deorowicz2021vcfshark, % VCFShark
dehaas2024genotype} %  GRG
most notably the Positional
Burrows--Wheeler Transform (PBWT)~\citep{durbin2014efficient}.
See~\citep{mcvean2019linkage} for a review of the techniques
employed in genetic data compression.
The widely-used PLINK binary format stores genotypes in a 
packed binary representation, supporting only biallelic
variants without phase information.
The PLINK 2 PGEN format~\citep{pgen2024} is more general
and compact than PLINK, compressing variant data using specialised
algorithms~\cite{sambo2014compression}.
Methods have also been developed which store variation data 
along with annotations in databases to facilitate
efficient queries~\cite[e.g.][]{
paila2013gemini,%GEMINI
lopez2017hgva} %OpenCGA
which either limit to certain classes of variant~\cite[e.g.][]{greene2023genetic}
or have storage requirements larger
than uncompressed VCF~\citep{al2023critical}. 
The SeqArray package~\citep{zheng2017seqarray} builds on the 
Genomic Data Storage container format~\cite{zheng2012high}
to store VCF genotype data in a packed and compressed format,
and is used in several downstream R packages~\cite[e.g.][]{
gogarten2019genetic,fernandes2020simplephenotypes}.

VCF is a row-wise format in which 
observations and metadata for a single variant are
encoded as a line of text~\citep{danecek2011variant}.
BCF~\citep{li2011statistical}, the standard binary representation of VCF,
is similarly row-wise, as 
are the majority of proposed alternative storage 
formats.
Row-wise storage makes retrieving all information
for a given record straightforward and efficient,
and works well when records are either relatively small
or we typically want to analyse each record in its entirety.
When we want to analyse only a subset of a record,
row-wise storage can be inefficient because we will usually need to
retrieve more information than required from storage. In the case 
of VCF (and BCF) where records are not of a fixed size and 
are almost always compressed in blocks, accessing any information
for a set of rows means retrieving and decompressing \emph{all} 
information from these rows.

The usual alternative to row-wise storage is \emph{columnar} storage:
instead of grouping together all the fields for a record,
we group together all the records for a given field.
Columnar storage formats such as Parquet~\citep{parquet2024}
make retrieving particular columns much 
more efficient and can lead to substantially better compression.
While columnar techniques have been successfully applied 
in alignment
% TODO check
% Jeff: OK with the ADAM reference in here? You did some parquet based
% alignment storage, right?
storage~\citep[e.g.][]{bonfield2014scramble,nothaft2015rethinking,bonfield2022cram},
the use of columnar technologies for
storing and analysing variation data have had limited
success~\citep{boufea2017managing,fan2020variant}.
Mapping VCF directly to a columnar layout, in which there is a 
column for the genotypes (and other per-call QC metrics) 
for each sample leads to a large number of columns, which 
can be cumbersome and cause scalability issues.
Fundamentally, columnar methods are one-dimensional, storing a vector
of values associated with a particular key, whereas 
genetic variation data is usually modelled as a two-dimensional matrix
in which we are interested in accessing both rows \emph{and} columns.
Just as row-oriented storage makes accessing data for a given
sample inefficient, columnar storage makes accessing all the data 
for a given variant inefficient.

\begin{figure}
\resizebox{225pt}{!}{\input{diagrams/data-model.tex}}
\caption{Chunked compressed storage of VCF data using Zarr. 
The \texttt{call\_genotype} array is a three-dimensional (variants, samples,
ploidy) array of integers, split into a uniform grid of 
chunks determined by the variant and sample chunk sizes (10,000
and 1,000 by default in \texttt{vcf2zarr}). Each chunk is associated 
with a key defining its location in this grid, which can be stored 
in any key-value store such as a standard file-system or cloud object
store. Chunks are compressed independently using standard 
codecs and pre-compression filters, which can be specified on a per-array
basis. Also shown are the one-dimensional \texttt{variant\_contig} (CHROM)
and \texttt{variant\_position} arrays (POS). Other fields are stored 
in a similar fashion. \label{fig-data-model}}
\end{figure}

VCF is at its core an encoding of the genotype matrix, where each entry
describes the observed genotypes for a given sample at a given variant site,
interleaved with per-variant information
and other call-level matrices (e.g., the GQ or AD fields).
The data is largely numerical and of fixed dimension, 
and is therefore a natural mapping to array-oriented 
or ``tensor'' storage.
We propose the VCF Zarr specification which maps the 
VCF data model into an array-oriented layout using Zarr 
(Fig~\ref{fig-data-model}). 
In the VCF Zarr specification, 
each field in a VCF is mapped to a separately-stored array, 
allowing for efficient retrieval and 
high levels of compression.
% In particular, call-level data is stored as $m \times n$ arrays
% (for $m$ sites and $n$ samples), allowing for efficient 
% retrieval of subsets of those fields along both the 
% variants and samples axis.
See the Methods for more detail on Zarr and the VCF Zarr
specification.

\begin{figure}
\begin{center}
\includegraphics[]{figures/data-scaling}
\end{center}
\caption{Compression performance on simulated genotypes.
Comparison of total stored bytes for VCF data produced 
by subsets of a large simulation of French-Canadians.
Sizes for $10^6$ samples are shown on the right. Sizes 
for Savvy (21.25GiB) and Zarr (22.06GiB) are very similar.
Also shown for reference is the size of genotype matrix 
when encoded as two bits per diploid genotype (2bit), as used 
in the PLINK binary format.
\label{fig-data-storage}}
\end{figure}

One of the key benefits of Zarr is its cloud-native design, 
but it also works well on standard file systems, where 
arrays and chunks are stored hierarchically in directories
and files (storage as a single Zip archive is also supported).
To enable comparison with the existing file-based ecosystem
of tools, we focus on Zarr's file system chunk storage in a series of illustrative 
benchmarks in the following sections. 
(See \citep{durbin2020task,moore2021ome,gowan2022using} for Zarr
benchmarks in cloud settings.)
We compare primarily with
VCF/BCF based workflows using \texttt{bcftools} because this
is the standard practice, used in the vast majority of cases.
We also compare with two representative recent specialised utilities;
see~\cite{danek2018gtc,zhang2023gbc} for further benchmarks of 
these and other tools.
Genozip~\cite{lan2020genozip,lan2021genozip} is a tool focused 
on compression performance, which uses a custom file format 
and a CLI to extract VCF as text with various filtering options.
Savvy~\cite{lefaive2021sparse} is an extension of BCF which 
takes advantage of sparsity in the genotype matrix as well
as using PBWT-based approaches for improved compression.
Savvy provides a CLI as well as a C++ API.
Our benchmarks are based on genotype data 
from subsets of a large and highly realistic 
simulation of French-Canadians~\cite{anderson2023on}
(see Methods for details on the dataset and benchmarking methodology).
Note that while simulations cannot capture 
all the subtleties of real data, the allele frequency
and population structure patterns in this dataset 
have been shown to closely follow 
observations~\cite{anderson2023on} and so it provides 
a reasonable and easily reproducible data point 
when comparing such methods.
The simulations only contain genotypes without any additional
high-entropy QC fields, which is unrealistic 
(see the Genomics England case-study
for benchmarks on a large human dataset that includes 
many such fields). 
Note, however, that such minimal, genotype-only data 
is something of a best-case scenario for specialised genotype
compression methods using row-wise storage.

Fig~\ref{fig-data-storage} shows compression performance 
on up to a million samples for chromosome 21, with 
the size of the genotype-matrix encoded as 1-bit per haploid
call included for reference.
% 2bit  1689.153258 G
% vcf   81.375831 G
% = 20.75 X, 4.8%
Gzip compressed VCF performs remarkably well, compressing 
the data to around 5\% of the 
minimal binary encoding of a biallelic genotype matrix
for 1 million samples. 
BCF provides a significant improvement in compression
performance over VCF (note the log-log scale). Genozip has 
superb compression, having far smaller file sizes that the 
other methods (although somewhat losing its advantage at 
larger sample sizes). Zarr and Savvy have 
almost identical compression performance in this example.
It is remarkable that the simple approach of compressing
two dimensional chunks of the genotype matrix 
using the Zstandard compressor~\citep{collet2021rfc} and the 
bit-shuffle filter from Blosc~\cite{alted2010modern} 
(see Methods for details) produces 
compression levels competitive with the highly specialised methods
used by Savvy.

\subsection{Calculating with the genotype matrix}
Storing genetic variation data compactly is important, but it is also
important that we can analyse the data efficiently. Bioinformatics 
workflows tend to emphasise text files and command line utilities 
that consume and produce text~\citep[e.g.][]{buffalo2015bioinformatics}. 
Thus, many tools that compress VCF data provide a command line 
utility with a query language to restrict the records
examined, perform some pre-specified calculations and finally 
output some text, typically VCF or tab/comma separated 
values~\citep{
layer2016efficient, %GQT
li2016bgt, % BGT
danek2018gtc, % GTC
lin2020sparse, % SpVCF
lan2020genozip,lan2021genozip, %genozip
zhang2023gbc}. %GBC
These pre-defined calculations are by necessity limited in scope, however,
and the volumes of text involved in Biobank scale datasets
make the classical approach of custom
analyses via Unix utilities in pipelines prohibitively slow. Thus, 
methods have begun to provide Application Programming Interfaces
(APIs), providing efficient access to genotype and other VCF 
data~\cite[e.g.][]{kelleher2013processing,lefaive2021sparse,
wertenbroek2022xsi}. By providing programmatic access,
the data can be retrieved from storage, decoded and then analysed
in the same memory space without additional copies and 
inter-process communication through pipes.

To demonstrate the accessibility of genotype data and efficiency with 
which calculations can be performed under the different formats,
we use the \texttt{bcftools +af-dist} plugin
(which computes a table of
deviations from Hardy-Weinberg expectations in 
allele frequency bins) as an example.
% The details of the \texttt{af-dist} operation are not important:
% as an example of a whole-matrix operation. 
We chose this particular operation for several reasons. 
First, it is a straightforward calculation that 
requires examining every element in the genotype matrix,
and can be reproduced in different programming languages
without too much effort.
Secondly, it produces a small volume of output  and therefore the 
time spent outputting results is negligible.
Finally, it has an efficient implementation written using the 
\texttt{htslib} C API~\citep{bonfield2021htslib},
and therefore running this command on a VCF or BCF file provides 
a reasonable approximation of the limit of what can be achieved in terms 
of whole-matrix computation on these formats.

\begin{figure}
\includegraphics{figures/whole-matrix-compute}
\caption{Whole-matrix compute performance with increasing sample size.
Total CPU time required to run \texttt{bcftools +af-dist}
and equivalent operations in a single thread for various tools.
Elapsed time is also reported (dotted line). Run-time for genozip
and bcftools on VCF
at $10^6$ samples were extrapolated by fitting an exponential.
See Methods for full details.
\label{fig-whole-matrix-compute}}
\end{figure}

Fig~\ref{fig-whole-matrix-compute} shows timing results 
for running \texttt{bcftools +af-dist} and equivalent operations 
on the data of Fig~\ref{fig-data-storage}. There is a large
difference in the time required (note the log-log scale). 
The slowest approach uses Genozip. Because Genozip does not
provide an API and only outputs VCF text, the best approach available 
is to pipe its output into \texttt{bcftools +af-dist}. 
This involves first decoding the data from Genozip format,
then generating large volumes  of VCF text (terabytes, in the 
largest examples here), which we must 
subsequently parse before finally doing the actual calculation. 
Running \texttt{bcftools +af-dist} directly on the gzipped VCF
is substantially faster, indicating that Genozip's excellent
compression performance comes at a substantial decompression cost.
Using a BCF file is again significantly faster,
because the packed binary format avoids the overhead of parsing 
VCF text into \texttt{htslib}'s internal data structures. 
We only use BCF for subsequent \texttt{bcftools} benchmarks.

The data shown in Fig~\ref{fig-whole-matrix-compute} for Zarr and Savvy
is based on custom programs written using their respective APIs
to implement the \texttt{af-dist} operation. The Zarr program uses
the Zarr-Python package to iterate over the decoded chunks of the 
genotype matrix and classifies genotypes within a chunk using a 14 line Python
function, accelerated using the Numba JIT compiler~\cite{lam2015numba}.
The allele frequencies and genotype counts are then analysed to produce 
the final counts within the allele frequency bins with 9 lines of 
Python using NumPy~\cite{harris2020array} functions. Remarkably, this 
short and simple Python program is substantially faster than the 
equivalent compiled C using \texttt{htslib} APIs on BCF (6.9 hours
vs 20.6 hours for 1 million samples). 
% num_samples  num_sites      tool  user_time  sys_time     wall_time  total_time
%     1000000    7254858     savvy    8371.07      4.16   8377.718136  2.326453
%     1000000    7254858  bcftools   73939.32     46.09  74023.083678  20.551503
%     1000000    7254858      zarr   24709.64     29.32  24750.704171  6.871933
The fastest method is the 
C++ program written using the Savvy API. This would largely seem
to be due to Savvy's excellent genotype decoding performance
% zarr 1.2 GiB
% savvy 6.6 GiB
% zarr_nshf 3.9 GiB
(up to 6.6GiB/s vs 1.2GiB/s for Zarr on this dataset;
Fig~\ref{fig-whole-matrix-decode}).
% sequence_length  num_samples  num_sites       tool       size
% 40         48129895      1000000    7254858       zarr  22.071687
% 35         48129895      1000000    7254858        bcf  51.749294
% 36         48129895      1000000    7254858    genozip  10.691384
% 37         48129895      1000000    7254858        sav  21.249436
% 38         48129895      1000000    7254858        tsk   1.802636
% 39         48129895      1000000    7254858        vcf  81.375831
% 41         48129895      1000000    7254858  zarr_nshf  29.897540
Turning off the BitShuffle filter for the Zarr dataset,
however, leads to a substantial increase in decoding speed
(3.9GiB/s) at the cost of a roughly 25\% increase in storage
space (29.9GiB up from 22.1GiB for 1 million samples; data not
shown). Given the relatively small contribution of genotypes to the
overall storage of real datasets (see the Genomics England example)
and the frequency that they are likely to be accessed, this
would seem like a good tradeoff in most cases.
This ability to easily tune compression performance
and decoding speed on a field-by-field basis is a major strong
point of Zarr. The \texttt{vcf2zarr} utility also provides
functionality to aid with such storage schema tuning.

\subsection{Subsetting the genotype matrix}
\begin{figure}[t]
\includegraphics{figures/subset-matrix-compute}
\caption{Compute performance on subsets of the matrix.
Total CPU time required to run the af-dist calculation for
a contiguous subset of 10,000 variants $\times$ 10 samples 
from the middle of the matrix
for the data in Fig~\ref{fig-data-storage}.
The \texttt{genozip} and \texttt{bcftools} pipelines involve
multiple commands required to correctly calculate the AF INFO field
required by \texttt{bcftools +af-dist}. See the Methods for full details
on the steps performed.
\label{fig-subset-matrix-compute}}
\end{figure}

As datasets grow ever larger, the ability to efficiently access subsets 
of the data becomes increasingly important. VCF/BCF achieve efficient 
access to the data for genomic ranges 
by compressing blocks of adjacent records using \texttt{bgzip},
and storing secondary indexes alongside the original 
files with a conventional suffix~\citep{li2011tabix}. 
Thus, for a given range query we 
decompress only the necessary blocks and can quickly access
the required records. 
The row-wise nature of VCF (and most proposed alternatives), however, means
that we cannot efficiently subset \emph{by sample}
(e.g., to calculate statistics within a particular cohort). In the extreme
case, if we want to access only the genotypes for a single sample
we must still retrieve and decompress the entire dataset.

We illustrate this cost of row-wise encoding in
Fig~\ref{fig-subset-matrix-compute}, where we run the af-dist calculation
on a small fixed-size subset of the genotype matrices of
Fig~\ref{fig-data-storage}. The two-dimensional chunking of Zarr
means that this sub-matrix can be efficiently
extracted, and therefore the execution time depends very weakly on 
the overall dataset size, with the computation requiring around
1 second for 1 million samples. Because of their 
row-wise encoding, CPU time scales with the number of samples
for all the other methods.
Fig~\ref{fig-subset-matrix-compute-supplemental} shows performance
for the same operation when selecting half of the samples in the 
dataset.

\subsection{Extracting, inserting and updating fields}
\begin{figure}
\includegraphics{figures/column-extract}
\caption{Time to extract the genome position and write to a text file.
Total CPU time required to extract the POS field for BCF,
sav and Zarr formats
for the data in Figure~\ref{fig-data-storage}.
For the BCF file we used \texttt{bcftools query -f"\%POS$\backslash$n"}.
For sav, we used the Savvy C++ API to extract position for each variant
and output text using the \texttt{std::cout} stream. For Zarr, we read 
the variant\_position array into a NumPy array, and then wrote to
a text file using the Pandas \texttt{write\_csv} method. 
Zarr CPU time is dominated by writing the text output; we also show
the time required to populate a NumPy array with the data in Zarr,
which is less than a second. Wall-clock time (dotted line) is dominated 
in this case by file I/O. Time to output text for Savvy is not significant
for $> 1000$ samples (not shown).
\label{fig-column-extract}}
\end{figure}
We have focused on the genotype matrix up to this point, contrasting
Zarr with existing row-wise methods.
Real-world VCFs encapsulate much more than just the genotype
matrix, and can contain large numbers of additional fields.
Fig~\ref{fig-column-extract} shows the time required to extract 
the genomic position of each variant in the simulated benchmark 
dataset, which we can use as an indicative example of a per-variant 
query. Although Savvy is many times faster than \texttt{bcftools query}
here, the row-wise storage strategy that they share means that 
the entire dataset must be read into memory and 
decompressed to extract just one field from each record. Zarr
excels at these tasks: we only read and decompress the information required.

Many of the additional fields that we find in real-world VCFs are 
variant-level annotations, extensively used in downstream applications.
For example, a common workflow is to 
add or update variant IDs in a VCF using a reference database
such as dbSNP \cite{Sherry2001dbSNP}. The standard approach to this
(using e.g.\ \texttt{bcftools annotate}) is to create a \emph{copy} of 
the VCF which includes these new annotations. Thus, even though 
we many only be altering a single field comprising a tiny fraction 
of the data, we still read, decompress, update, compress and 
write the entire dataset to a new file. With Zarr,
we can update an existing field or add arbitrary additional
fields without touching the rest of the data or creating redundant
copies.

\subsection{Case study: Genomics England 100,000 genomes}
In this section we demonstrate the utility of VCF Zarr on a large human dataset
and the scalability of the \texttt{vcf2zarr} conversion utility.
Genomics England’s multi-sample VCF dataset (aggV2) is an 
aggregate of 78,195 gVCFs from rare disease and cancer participants 
recruited as part of the 100,000 Genomes Project~\cite{turnbull2018100}. 
The dataset comprises approximately 722 million annotated single-nucleotide 
variants and small indels split into 1,371 roughly equal chunks and 
totalling 165.3 TiB of VCF data after \texttt{bgzip} compression. 
The dataset is used for a variety of research purposes, ranging from 
GWAS~\cite{kousathanas2022whole} and 
imputation~\cite{shi2023genomics} to 
simple queries involving single gene 
regions~\cite{leggatt2023genotype,lam2023repeat}.

As described in the Methods, conversion to Zarr using 
\texttt{vcf2zarr} is a two-step process. We 
first converted the 106 VCF files (12.81 TiB) for chromosome 2
into the intermediate columnar format (ICF). This task was 
split into 14,605 partitions, and distributed using the Genomics England
HPC cluster. The average run-time per partition was 20.7 min.
The ICF representation used a total
of 9.94 TiB over 3,960,177 data storage files. 
We then converted the ICF to Zarr, partitioned into
5989 independent jobs, with an 18.6 min average run time.
This produced a dataset with 44 arrays, consuming a 
total of 2.54 TiB of storage over 6,312,488
chunk files. This is a roughly 5X reduction in total storage 
space over the original VCF. 
The top fields in terms 
of storage are detailed in Table~\ref{tab-genomics-england-data}.
We do not compare with other tools such as Genozip and 
Savvy here because  they have fundamental limitations 
(as shown in earlier simulation-based benchmarks),
and conversion of these large VCFs is a major undertaking.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced for 
Genomics England aggV2 VCFs on chromosome 2 using \texttt{vcf2zarr}
default settings. Each field is stored independently 
as a Zarr array with the given type (sufficient to represent all values in the
data). We show the total storage consumed (reported via \texttt{du}) in 
power-of-two units, and the compression ratio achieved on that array.
We also show the percentage of the overall storage that each array consumes
(omitting values < 0.01\%).
\label{tab-genomics-england-data}}
\begin{tabular}{llS[table-format=3.1]S[table-format=3.2]S[table-format=3.2]}
\toprule
{Field} & {type} & {storage} & {compress} & {\%total} \\
\midrule
/call\_AD & int16 & 658.4G & 26 & 25.35\% \\
/call\_GQ & int16 & 654.5G & 13 & 25.20\% \\
/call\_DP & int16 & 570.0G & 15 & 21.95\% \\
/call\_DPF & int16 & 447.1G & 20 & 17.22\% \\
/call\_PL & int16 & 162.6G & 160 & 6.26\% \\
/call\_GQX & int16 & 41.0G & 210 & 1.58\% \\
/call\_FT & string & 25.0G & 1400& 0.96\% \\
/call\_genotype & int8 & 21.5G & 410& 0.83\% \\
/call\_genotype\_mask & bool & 12.8G & 680& 0.49\% \\
/call\_genotype\_phased & bool & 2.4G & 1900& 0.09\% \\
/call\_PS & int8 & 383.4M & 12000& 0.01\% \\
% /call\_ADF & int8 & 383.4M & 12000.0 & 0.01\% \\
% /call\_ADR & int8 & 383.4M & 12000.0 & 0.01\% \\
/variant\_position & int32 & 111.6M & 2 & \\
/variant\_quality & float32 & 87.4M & 2.6 & \\
/variant\_allele & string & 69.3M & 13 & \\
% /variant\_OLD\_MULTIALLELIC & object & 55.88M & 8.2 & 0.00\% \\
/variant\_AN & int32 & 47.3M & 4.8 & \\
% /variant\_AC & int32 & 47M & 4.9 & 0.00\% \\
% /variant\_AC\_Het & int32 & 46.62M & 4.9 & 0.00\% \\
% /variant\_ABratio & float32 & 20.23M & 11.0 & 0.00\% \\
% /variant\_MendelSite & float32 & 20.09M & 11.0 & 0.00\% \\
% /variant\_medianGQ & float32 & 19.16M & 12.0 & 0.00\% \\
% /variant\_completeGTRatio & float32 & 14.98M & 15.0 & 0.00\% \\
% /variant\_phwe\_eur & float32 & 14.46M & 16.0 & 0.00\% \\
% /variant\_phwe\_afr & float32 & 13.37M & 17.0 & 0.00\% \\
% /variant\_medianDepthNonMiss & float32 & 12.17M & 19.0 & 0.00\% \\
% /variant\_medianDepthAll & float32 & 11.99M & 19.0 & 0.00\% \\
% /variant\_phwe\_sas & float32 & 11.59M & 20.0 & 0.00\% \\
% /variant\_AC\_Hom & int32 & 11.12M & 21.0 & 0.00\% \\
% /variant\_missingness & float32 & 9.3M & 25.0 & 0.00\% \\
% /variant\_phwe\_eas & float32 & 7.12M & 32.0 & 0.00\% \\
/variant\_filter & bool & 6.4M & 570 & \\
% /variant\_AC\_Hemi & int32 & 849.43 KiB & 280.0 & 0.00\% \\
% /variant\_composite\_filter & bool & 433.38 KiB & 130.0 & 0.00\% \\
% /variant\_OLD\_CLUMPED & object & 384.18 KiB & 1200.0 & 0.00\% \\
% /variant\_id & object & 298.76 KiB & 1600.0 & 0.00\% \\
% /variant\_phwe\_amr & float32 & 269.52 KiB & 870.0 & 0.00\% \\
% /variant\_id\_mask & bool & 269.46 KiB & 220.0 & 0.00\% \\
/sample\_id & str & 268.1K & 2.3 & \\
% /variant\_contig & int16 & 257.77 KiB & 450.0 & 0.00\% \\
% /contig\_length & int64 & 3.79 KiB & 5.3 & 0.00\% \\
% /contig\_id & object & 2.3 KiB & 8.8 & 0.00\% \\
% /filter\_id & object & 808 bytes & 0.6 & 0.00\% \\
%%%% OLD for chr20
% call\_AD &  int16 & 179.7G & 26 & 24.0\\
% call\_GQ &  int16 & 171.8G & 13 & 23.0 \\
% call\_DP &  int16 & 141.8G & 16 & 18.9 \\
% call\_DPF& int16  & 115.1G & 20 & 15.3\\
% call\_FT &  string & 58.5G & 160 & 7.8 \\
% call\_PL &  int16 & 51.4G & 140 & 6.9 \\
% call\_GQX &  int16 & 12.1G & 190 & 1.6 \\
% call\_genotype & int8 & 6.1G & 380 & 0.8 \\
% call\_genotype\_mask & bool & 3.7G  & 630 & 0.5\\
% call\_genotype\_phased & bool & 692.5M  & 1700 & 0.1 \\
% call\_PS  & int8  & 102.2M & 12000 & 0.05 \\
% variant\_quality & float32 & 22.6M & 2.7 & <0.01 \\
% variant\_allele & string & 21.5M & 11 \\
% variant\_position & int32 & 12.7M & 4.7 \\
% variant\_ABratio & float32 & 9.6M & 6.3 \\
% variant\_AN & int32 & 9.6M & 6.3 \\
% variant\_filter & bool & 1.9M & 490 \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab-genomics-england-data} shows that the dataset storage
size is dominated by a few columns with the top four
(call\_AD, call\_GQ, call\_DP and call\_DPF) accounting for 
90\% of the total. These fields are much less compressible
than genotype data (which uses $<1\%$ of the total space here)
because of their inherent noisiness~\citep{lin2020sparse}. 
Note that these top four fields are stored as 16 bit integers
because they contain rare outliers that cannot be stored as 
8 bits. While the fields could likely be truncated to have 
a maximum of 127 with minimal loss of information, 
the compression gains from doing so are relatively minor,
and we therefore opt for fully lossless compression here 
for simplicity. The call\_PS field here has an extremely high
compression ratio because it consists entirely of missing data
(i.e., it was listed in the header but never used in the VCF).

To demonstrate the computational accessibility of Zarr on this 
large human dataset, we performed some illustrative benchmarks.
As these benchmarks take some time to run, we focus 
on a single 132GiB compressed VCF file covering
positions 58,219,159--60,650,943 (562,640 variants) 
from the middle of the list of 106 files for chromosome 2.
We report both the total CPU time and elapsed wall-clock time here
as both are relevant.
First, we extracted the genome position for each variant in this single VCF 
chunk using \texttt{bcftools query} and Python Zarr code as described in 
Fig~\ref{fig-column-extract}. The \texttt{bcftools} command required 
% def parse_time(time_str):
%     minsplit = time_str.split("m")
%     return datetime.timedelta(minutes=int(minsplit[0]),
%         seconds=float(minsplit[1][:-1]))
% def parse_time_output(s):
%     values = {}
%     for line in s.splitlines():
%         split = line.split()
%         values[split[0]] = parse_time(split[1])
%     return values["real"].seconds / 60, (values["user"] + values["sys"]).seconds / 60
% real  85m51.013s
% user  54m39.986s
% sys 0m45.277s
55.42 min CPU and 85.85 min elapsed.
% CPU times: user 846 ms, sys: 1.93 s, total: 2.78 s
% Wall time: 1min 44s
% 55.42 * 60 / 2.78  = 1196
% 85.85 / 1.73 = 49.62 
The Zarr code required 2.78 sec CPU and 1.73 min elapsed.
This is a 1196X smaller CPU burden and a 50X speed-up in elapsed time.
The major difference between CPU time and wall-time is noteworthy 
here, and indicates some opportunities for improvement in VCF Zarr
in high-latency environments such as the shared file system in the
Genomics England HPC system. Currently VCF Zarr does not store any
specialised index to map genomic coordinates to array positions
along the variants dimension. Instead, to find the relevant slice
of records corresponding to the range of positions in the target
VCF file, we load the entire variant\_position array and 
binary search. This entails reading 5,989 chunk files
(the chunk size is 100,000 variants) which incurs a substantial
latency penalty on this system. Later versions of the specification
may solve this problem by storing an array of size 
(approximately) the number variant chunks 
which maps ranges of genome coordinates to chunk indexes,
or a more specialised structure that supports overlap queries.

% When we subsequently time the operation of 
% just populating an array of the variant positions in specified region,
% this is reduced to 1.13 sec, with an elapsed time of 20.9 sec.
% [Note: this is crazy that we still have a 19 second I/O wait. The
% 5989 files for the POS chunks must be in cache already, so the delay
% has to be the file system checking cache coherency.]

We then ran the af-dist calculation (Figs~\ref{fig-whole-matrix-compute}
and~\ref{fig-subset-matrix-compute}) on the VCF file
using \texttt{bcftools +af-dist} as before. 
The elapsed time for this operation was 716.28 min CPU, 716.3 min elapsed.
Repeating this operation for the same coordinates in Zarr 
(using Python code described in previous sections) 
% 716.28 / 2.32 = 308.74
% 716.3 / 4.25 = 168.54
gave a total CPU time of 2.32 min and elapsed time of 4.25 min.
This is a 309X reduction in CPU burden and a 169X speed-up in elapsed time.
It is worth noting here that \texttt{bcftools +af-dist} cannot be 
performed in parallel across multiple slices of a chromosome, and if 
we did want to run it on all of chromosome 2 we would need to 
concatenate the 106 VCF files. While af-dist itself is not a common operation,
many tasks share this property
of not being straightforwardly decomposable across multiple VCF files.

Finally, to illustrate performance on a common filtering task, we 
created a copy of the VCF chunk which contains only variants 
that pass some common filtering criteria using
\texttt{bcftools view -I --include "FORMAT/DP>10 \& FORMAT/GQ>20"},
following standard practices~\citep[e.g.][]{bergstrom2020insights,
kousathanas2022whole,
chen2024genomic}.
This used 689.46 min CPU time, with an elapsed time of 689.48 min.
In comparison, computing and storing a variant mask (i.e., a boolean value 
for each variant denoting whether it should be considered or not
for analysis) based on the same criteria using Zarr 
consumed 1.96 min CPU time with an elapsed time of 11 min.
This is a 358X reduction in CPU usage, and 63X reduction in elapsed time.
% 689.46 / 1.96 = 351.76
% 689.48 / 11 = 62.6 = 63X
There is an important distinction here between creating a copy
of the data (an implicit part of VCF based workflows) and creating an
additional \emph{mask}. As Table~\ref{tab-genomics-england-data} 
illustrates, call-level masks are cheap (the standard 
genotype missingness mask, call\_genotype\_mask, uses 0.49\% of the overall 
storage) and variant or sample level masks require negligible storage.
If downstream software can 
use configurable masks (at variant, sample and call level) 
rather than expecting full copies of the data, major storage savings
and improvements in processing efficiency can be made.
The transition from the manifold inefficiencies of 
present-day ``copy-oriented'' computing,
to the ``mask-oriented'' analysis of large immutable, single-source
datasets is a potentially transformational change enabled by Zarr.

\section{Discussion}
% Zarr is great
VCF is a central element of modern genomics, facilitating
the exchange of data in a large ecosystem of interoperating tools. 
Its current row-oriented form, however,
is fundamentally inefficient,
profoundly limiting the scalability of the present generation
of bioinformatics tools. Large scale VCF data cannot 
currently be 
processed without incurring a substantial economic
(and environmental~\cite{grealey2022carbon}) cost.
We have shown here that this is not a necessary situation,
and that greatly improved efficiency can be achieved by 
using more appropriate storage representations tuned
to the realities of modern computing. We have argued that 
Zarr provides a powerful basis for cloud-based
storage and analysis of large-scale genetic variation data.
We propose the VCF Zarr specification which losslessly
maps VCF data to Zarr, and provide an efficient and scalable
tool to perform conversion.

% We're not pretending is a solution to all problems and
% up front about weaknesses of the model.
Zarr provides pragmatic solutions to some of the more pressing 
problems facing the analysis of large-scale genetic variation
data, but it is not a panacea. Firstly, 
any dataset containing a variant with a large number of alleles
(perhaps due to indels) will cause problems because the 
dimensions of fields are determined by their \emph{maximum}
dimension among all variants. In particular this is problematic
for fields like PL in which the dimension depends 
quadratically on the number of alleles (although practical
solutions have been suggested that we plan to 
implement~\cite{poterba2024scalable}).
Secondly, the design of 
VCF Zarr emphasises efficiency of analysis for a fixed 
dataset, and does not consider how samples (and the 
corresponding novel variants) should be added.
Thirdly, Zarr works best for numerical data of a fixed 
dimension, and therefore may not suitable for representing
the unstructured data often included in VCF INFO fields.

Nonetheless, there are numerous datasets that exist today
that would likely reap significant benefits from being deployed 
in a cloud-native fashion using Zarr. Object 
stores typically allow for individual objects (chunks, in
Zarr) to be associated with ``tags'', which can then be 
used to associate storage class, user access control
and encryption keys. 
Aside from the performance benefits
we have focused on here provided by Zarr, the ability
to (for example) use high-performance storage for commonly
used data such as the variant position and 
more cost-effective storage classes
for infrequently used bulk QC data should provide 
significant operational benefits.
Granular access controls would similarly allow non-identifiable
variant-level data to be shared relatively freely,
with genotype and other data more tightly controlled
as required.
Even finer granularity is possible if samples are grouped by
access level within chunks (padding partially filled 
chunks as needed and using an appropriate sample mask). 
Providing client applications direct access to 
the data over HTTP
and delegating access control to the cloud provider
makes custom web APIs~\cite{kelleher2019htsget} 
and cryptographic container formats~\citep{senf2021crypt4gh}
largely unnecessary.

The VCF Zarr specification and scalable \texttt{vcf2zarr} conversion utility
provided here are a necessary starting point for such cloud-native
biobank repositories and open up many possibilities,
but significant investment and development would be needed 
to provide a viable alternative to standard bioinformatics workflows. 
Two initial directions for development, however, may quickly
yield sufficient results to both greatly improve researcher productivity on 
large, centrally managed datasets such as Genomics England
and motivate further research and development.
The first direction is to provide compatibility with existing 
workflows via a ``\texttt{vcztools}'' command line utility which 
implements a subset of \texttt{bcftools} functionality (such as 
\texttt{view} and \texttt{query}) on a VCF Zarr dataset.
Such a tool would speed up some common queries by orders 
of magnitude, and reduce the need for user orchestration of 
operations among manually split VCF chunks (large VCF datasets
are typically split into hundreds of files; see the Genomics
England case study). Datasets could then be hosted in cloud
object stores, while still presenting file-like semantics
for existing workflows. This could provide an evolutionary path,
allowing established analysis workflows to coexist with new Zarr-native
approaches, working from the same primary data.

The second natural direction for development is to create
these Zarr-native applications, which can take advantage
of the efficient data representation
across multiple programming languages (see Methods).
The Python data science ecosystem, in particular,
has a rich suite of powerful tools~\cite[e.g.][]{
mckinney2010data,
lam2015numba,
kluyver2016jupyter,
harris2020array,
virtanen2020scipy} 
and is increasingly
popular in recent biological applications~\citep[e.g.][]{
abdennur2020cooler,
rand2022bionumpy,
open2c2024bioframe,
hou2024admix}.
Dask~\cite{rocklin2015dask} and 
Xarray~\citep{hoyer2017xarray} provide powerful abstractions
for large-scale data analysis and can be
used to perform interactive analysis
of large VCF Zarr datasets,
with computations transparently distributed over a cluster.
As shown in our benchmarks, lower-level computations
in which optimised kernels are applied on a chunk-by-chunk
basis are also straightforward to implement and highly performant.
Thus, a range of possibilities exist in which developers can 
choose the appropriate level of abstraction and tool chain
on a case-by-case basis. These abstractions will extend, hopefully, to
include more convenient primitives for genomics workflows through
projects like sgkit in the future.

While Zarr is now widely used across the sciences (see Methods)
it was originally 
developed to store genetic variation data from the 
\textit{Anopheles gambiae} 
1000 Genomes Project~\citep{anopheles2017genetic}
and is in active use in this 
setting~\cite[e.g.][]{ahouidi2021open,trimarsanto2022molecular}.
The VCF Zarr specification presented here builds on this real-world 
experience but is still a draft proposal that would benefit 
from wider input across a range of applications. With some 
refinements and sufficient uptake it may be suitable 
for standardisation~\cite{rehm2021ga4gh}.
The benefits of Zarr are substantial, and, in certain settings,
worth the cost of retooling away from classical file-oriented workflows.
For example, the MalariaGEN Vector Observatory
currently uses Zarr to store data from whole-genome sequencing 
of 23,000 \textit{Anopheles} mosquitoes from 31 African 
countries~\cite{mvgo2024}.
The data is hosted in Google Cloud Storage and can be analysed interactively
using free cloud computing services like Google Colab,
enabling the use of data by scientists in malaria-endemic countries
where access to local computing infrastructure and sufficient network 
bandwidth to download large datasets may be limited.
VCF Zarr could similarly reduce the costs of analysing large-scale
human data, and effectively open access to biobanks
for a much broader group of researchers than currently possible.

\section{Methods}

\subsection{Zarr and block-based compression}
% Zarr is a simple layer on top of best-in-class, modern components
In the interest of completeness it is useful to provide a high-level overview
of Zarr and the technologies that it depends upon. Zarr is a specialised format
for storing large-scale $n$-dimensional data (arrays). Arrays
are split into chunks, which are compressed and stored separately. Chunks are 
addressed by their indexes along the dimensions of the array, and the 
compressed data associated with this key. Chunks can
be stored in individual files (as we do here), but a wide array of different
storage backends are supported including cloud object stores 
and NoSQL databases;
in principle, Zarr can store data in any key-value store.
Metadata describing the array and its properties is then stored 
in JSON format along with the chunks. The simplicity and transparency
of this design has substantial advantages over other technologies
such as HDF5~\citep{folk2011overview} which are relatively complex and opaque.
This simplicity has led to numerous implementations of the Zarr specification
being developed, ranging from the mature Zarr-Python~\citep{zarrpython}
and TensorStore~\citep{tensorstore} implementations
to more experimental extensions to packages like
GDAL~\citep{gdal_zarr},
NetCDF~\citep{netcfd_c},
N5~\citep{n5zarr}
and xtensor~\citep{xtensor_zarr}
as well as  standalone libraries for JavaScript~\cite{zarrjs},
Julia~\cite{zarrjl}, Rust~\citep{zarrs}
and R~\cite{pizzarr}.

Zarr is flexible in allowing different compression codecs and 
pre-compression filters to be specified on a per-array basis
(see section XXX).
Two key technologies often used in conjunction with Zarr are the Blosc
% This seems like the best citation for Blosc? It is mentioned here
meta-compressor~\cite{alted2010modern}
and Zstandard compression algorithm~\citep{collet2021rfc}.
Blosc is a high-performance compressor optimised for numerical
data which uses ``blocking''~\citep{alted2010modern} to 
optimise CPU-cache access patterns, as well as highly optimised
bit and byte shuffle filters.  Remarkably, on highly 
compressible datasets, Blosc decompression can be faster 
than \texttt{memcpy}.
% Mentioning this detail to reassure readers who might be thinking
% "how would this be implemented and what kind of horrible dependencies
% would I need?"
Blosc is written in C, with APIs for C, Python, Julia, Rust
and others.
Blosc is a ``meta-compressor'' because it provides 
access to several different compression codecs. The 
Zstandard codec is of particular 
interest here as it achieves very high compression ratios
with good decompression speeds (Figs~\ref{fig-whole-matrix-decode},
\ref{fig-compression-compressor}).
% This may seem a bit off topic, but want to reassure reader that 
% this isn't some niche technology
Zstandard is also used in several recent VCF compression 
methods~\citep[e.g.][]{lefaive2021sparse,wertenbroek2022xsi}.
% LZ4 is also notable - any citations for this?

Scientific datasets are increasingly overwhelming the classical
model of downloading and analysing locally, and are migrating to 
centralised cloud repositories \citep{abernathey2021cloud,moore2021ome}.
The combination of Zarr's simple and cloud-friendly storage 
of data chunks with state-of-the-art compression methods has 
led to Zarr gaining significant traction in these settings.
Multiple petabyte-scale datasets are now stored using 
Zarr~\cite[e.g.][]{gowan2022using, % weather model data
fahnestock2023mappin, % ITS_LIVE dataset
cmip6_dataset}
or under active consideration for migration~\citep{durbin2020task,abernathey2021opening}.
The Open GeoSpatial consortium has formally recognised Zarr as a community
standard~\cite{ogc_zarr2_standard}
and has formed
a new GeoZarr Standards Working Group to establish a Zarr encoding for
geospatial data~\cite{ogc_geozarr_news}.

Zarr has recently been gaining popularity in biological applications.
The Open Microscopy Environment has developed OME-Zarr~\cite{moore2023ome} as one 
of its ``next generation'' cloud ready file formats~\citep{moore2021ome}.
OME-Zarr already has a rich suite of supporting
tools~\cite{moore2023ome,rzepka2023toward}. 
Zarr has also seen recent uptake 
in single-cell single-cell genomics~\citep{dhapola2022scarf,virshup2023scverse}
and multimodal spatial omics
data~\citep{marconato2024spatialdata,baker2023emobject}.
Recent additions using Zarr include the application of deep learning
models to genomic sequence data~\citep{klie2023predictive}, storage and 
manipulation of large-scale linkage disequilibrium matrices \cite{zabad2023fast},  
and a browser for genetic variation data~\citep{konig2023divbrowse}.

\subsection{The VCF Zarr specification}
The VCF Zarr specification is a direct mapping from the VCF data model
to a chunked binary array format using Zarr,
and is an evolution of the Zarr format used in the \texttt{scikit-allel}
package~\citep{miles2023scikit}.
VCF Zarr takes advantage
of Zarr's hierarchical structure by representing a VCF file as a top-level
Zarr group containing Zarr arrays. Each VCF field (fixed fields, INFO fields,
and FORMAT fields) is represented as a separate array in the Zarr hierarchy.
Some of the structures from the VCF header are also represented as arrays,
including contigs, filters, and samples.

The specification defines the name, shape, dimension names, and data type
for each array in the Zarr store. These ``logical'' properties are mandated,
in contrast to ``physical'' Zarr array properties such as chunk sizes and
compression, which can be freely chosen by the implementation. This
separation makes it straightforward for tools and applications to consume
VCF Zarr data since the data has a well-defined structure, while allowing
implementations enough room to optimise chunk sizes and compression
according to the application's needs.

The specification defines a clear mapping of VCF field names (keys) to
array names, VCF Number to array shape, and VCF Type to array data type.
To take one example, consider the VCF AD genotype field defined by the
following VCF header: \texttt{
\#\#FORMAT=<ID=AD,Number=A,Type=Integer,Description="Allele Depths">}.
The FORMAT key \texttt{ID} maps to an array name of \texttt{call\_AD}
(FORMAT fields have a \texttt{call\_} prefix, while INFO fields have a
\texttt{variant\_} prefix; both are followed by the key name). Arrays
corresponding to FORMAT fields are 3-dimensional with shapes that look
like \texttt{(variants, samples, <Number>)} in general. In this case, the
Number A entry indicates that the field has one value per alternate allele,
which in VCF Zarr is represented as the \texttt{alt\_alleles} dimension name,
so the shape of this array is \texttt{(variants, samples, alt\_alleles)}.
The VCF Integer type can be represented as any Zarr integer type, 
and the specification doesn't mandate particular integer widths. 
The \texttt{vcf2zarr} (see the next section) conversion utility 
chooses the narrowest integer width that can represent the data in
each field.

An important aspect of VCF Zarr is that field dimensions are 
global and fixed, and defined as the maximum across all rows.
Continuing the example above, the third dimension of the array 
is the maximum number of alternate alleles across \emph{all} 
variants. For variants at which there are less than the maximum
number of alternative alleles, the third dimension of the 
\texttt{call\_AD} array
is padded with a sentinel value (-2 for integers and a specific
non-signalling NaN for floats). While this is not a problem 
in practice for datasets in which all four bases are observed, 
it is a substantial issue for
fields that have a quadratic dependency on the number of alleles
(Number=G) such as PL. Such fields are already known to cause 
significant problems, and the ``local alleles'' proposal 
provides an elegant solution~\cite{poterba2024scalable}.
As this approach is on a likely path to 
standardisation~\cite{danecek2021twelve}, we plan to 
include support in later versions of VCF Zarr.

The VCF Zarr specification can represent anything described by BCF
(which is somewhat more restrictive than VCF) except for two corner
cases related to the encoding of missing data. Firstly, VCF Zarr does
not distinguish between a field that is not present and one that 
is present but contains missing data. For example, a variant with an
INFO field \texttt{NS=.} is represented in the same way in VCF Zarr
as an INFO field with no \texttt{NS} key. Secondly, because of the use
of sentinel values to represent missing and fill values for integers
(-1 and -2, respectively), a field containing these original values
cannot be stored. In practice this doesn't seem to be much of 
an issue (we have not found a real VCF that contains negative 
integers). However, if -1 and -2 need to be stored, a float field
can be used without issues.

The VCF Zarr specification is general and can be mapped to 
file formats such as \toolname{plink}~\citep{purcell2007plink,chang2015second}
and BGEN~\citep{band2018bgen} with some minor extensions.

\subsection{vcf2zarr}
Converting VCF to Zarr at Biobank scale is challenging. 
One problem is to determine the dimension of fields,
(i.e., finding the maximum number of alternate alleles and the 
maximum size of \texttt{Number=.} fields) which requires a full
pass through the data. Another challenge is to keep
memory usage within reasonable limits: although 
we can view each record in the VCF one-by-one,
we must buffer a full chunk (10,000 variants is the default in
\texttt{vcf2zarr}) 
in the variants dimension for each of the fields to convert to Zarr. 
For VCFs with many FORMAT fields and large numbers of samples this can
require tens of gigabytes of RAM per worker, making
parallelism difficult. Reading the VCF multiple times for different fields
is possible, but would be prohibitively slow for multi-terabyte VCFs.

The \texttt{vcf2zarr} utility solves this problem by first converting 
the VCF data (which can be split across many files) into an Intermediate
Columnar Format (ICF). The \texttt{vcf2zarr explode} command takes a set
of VCFs, and reads through them using
\texttt{cyvcf2}~\cite{pedersen2017cyvcf2},
storing each field independently in (approximately) fixed-size 
compressed chunks.
Large files can be partitioned based on information extracted from the 
CSI or Tabix indexes, and so different parts of a file can be 
converted to ICF in parallel.
Once all partitions have completed, information about 
the number of records in each partition and chunk of a given
field is stored so that the record at a particular index 
can be efficiently retrieved. 
Summaries such as maximum dimension and the minimum and maximum value 
of each field are also maintained, to aid choice of data types later.
A set of VCF files can be converted to intermediate columnar 
format in parallel on a single machine 
using the \texttt{explode} command,
or can be distributed across a cluster using the 
\texttt{dexplode-init},
\texttt{dexplode-partition} and \texttt{dexplode-finalise} commands.

Once the VCF data has been converted to the intermediate columnar format,
it can then be converted to Zarr using the \texttt{vcf2zarr encode}
command. By default we choose integer widths based on the maximum
and minimum values observed during conversion to ICF
along with reasonable compressor defaults (see next section).
Default choices can be 
modified by generating a JSON-formatted storage schema,
which can be edited and supplied as an argument to \texttt{encode}.
Encoding a given field (for example, \texttt{call\_AD})
involves creating a buffer to hold a full variant-chunk of the 
array in question, and then sequentially filling this buffer with 
values read from ICF and flushing to file.
 Similar to the \texttt{explode} command,
encoding to Zarr can be done in parallel on a single 
machine using the \texttt{encode} command,
or can be distributed across a cluster using the 
\texttt{dencode-init},
\texttt{dencode-partition} and \texttt{dencode-finalise} commands.
The distributed commands are fault-tolerant, reporting any failed
partitions so that they can be retried.

\subsection{Choosing default compressor settings}
To inform the choice of compression settings across different fields 
in VCF data, we analysed their effect on compression ratio on 
recent high-coverage WGS data from the 1000 Genomes 
project~\citep{byrska2022high}. We began by downloading 
the first 100,000 lines of the VCF 
for chromosome 22 (giving a 1.1GiB compressed VCF)
and converted to Zarr using 
\texttt{vcf2zarr} with default settings. We then systematically 
examined the effects of varying chunk sizes and compressor settings
on the compression ratio for call-level fields. We excluded call\_PL
from this analysis as it requires conversion to a ``local alleles''
encoding~\citep{poterba2024scalable} to be efficient, 
which is planned for implementation in a future 
version of \texttt{vcf2zarr}.

Fig~\ref{fig-compression-compressor} shows the effect of varying compression
codecs in Blosc. The combination of outstanding compression performance
and competitive decoding speed (Fig~\ref{fig-whole-matrix-decode}) 
makes zstd a good default choice.

The \texttt{shuffle} parameter in the Blosc meta-compressor~\cite{alted2010modern}
can result in substantially better compression, albeit at the 
cost of somewhat slower decoding (see Fig~\ref{fig-whole-matrix-decode}).
Fig~\ref{fig-compression-shuffle} shows the effect of 
bit shuffle (grouping together bits at the same position
across bytes before compression), and byte shuffle (grouping together 
bytes at the sample position across words before compression)
on compression ratio. Bit shuffle provides a significant
improvement in compression for the call\_genotype field  because 
the vast majority of genotype calls will be 0 or 1, and therefore 
bits 1 to 7 will be 0. Thus, grouping these bits together will 
lead to significantly better compression. This strategy also 
works well when compressing boolean fields stored as 
8 bit integers, where the top 7 bits are always 0. In practice,
boolean fields stored in this way have very similar compression
to using a bit-packing pre-compression filter (data not shown).
Although byte shuffle leads to somewhat better compression for call\_AD
and call\_DP, it gives substantially worse compression on call\_AB
than no shuffling. 
The default in \texttt{vcf2zarr}
is therefore to use bit shuffle for \texttt{call\_genotype}
and all boolean fields, and to not use 
byte shuffling on any field. These defaults can be easily
overruled, however, by outputting and modifying a JSON formatted 
storage schema before encoding to Zarr.

Fig~\ref{fig-compression-chunksize} shows that chunk size has a 
weak influence on compression ratio for most fields. 
Increasing sample chunk size slightly increases 
compression on call\_AB, and has no effect on less compressible
fields. Variant chunk size appears to have almost no effect 
on compression ratio. 
Interestingly, the choice of chunk size along the sample dimension for the
genotype matrix does have a significant effect. 
With six evenly spaced points between 100 and 2504,
Fig~\ref{fig-compression-chunksize}A shows a somewhat unpredictable 
relationship between sample chunk size and compression ratio.
The more fine-grained analysis of
Fig~\ref{fig-compression-chunksize-finegrained} shows that three
distinct trend lines emerge depending on the chunk size 
divisibility, with the modulus 
(i.e., the remainder in the last chunk) also having a minor effect. 
At greater than 40X,
compression ratio is 
high in all cases, and given that genotypes contribute relatively
little to the total storage of real datasets
(Table~\ref{tab-genomics-england-data}) the effect will
likely be fairly minor in practice.
Thus, we do not expect the choice of chunk size to have a significant
impact on overall storage usage, and so choice may be
determined by other considerations such as expected data
access patterns.

\subsection{Benchmarks}
In this section we describe the methodology used for the simulation-based
benchmarks of Figs~\ref{fig-data-storage},\ref{fig-whole-matrix-compute},
\ref{fig-subset-matrix-compute} and \ref{fig-column-extract}.
The benchmarks use data simulated by conditioning on a large
pedigree of French-Canadians using 
\texttt{msprime}~\citep{baumdicker2021efficient},
which have been shown to follow patterns observed in real 
data from the same population to a remarkable 
degree~\cite{anderson2023on}.
We begin by downloading the simulated ancestral recombination 
graph~\cite{brandt2024promise,lewanski2024era,wong2024general}
for chromosome 21 from Zenodo~\cite{anderson2023simulated} 
in compressed \texttt{tszip} format. This 552M file 
contains the simulated ancestry and mutations for 1.4 million
present-day samples. We then subset the full simulation 
down to ${10^1, 10^2, \dots, 10^6}$ samples using 
ARG simplification~\cite{kelleher2018efficient,wong2024general},
storing the subsets in \texttt{tskit} format~\cite{tskit2024}.
Note that this procedure captures the growth 
in the number of variants (shown in the top x-axis labels)
as we increase sample sizes as a 
natural consequence of population-genetic processes.
% jk@holly$ python3 src/collect_data.py site-allele-report
% scaling/data/chr21_10_6.zarr/
% 0 alleles:  100.00%
% 1 alleles:  100.00%
% 2 alleles:  7.92%
% 3 alleles:  0.21%
As a result of simulated mutational processes,
most sites have one alternate allele, 
with 7.9\% having two and 0.2\% having three
alternate alleles in the $10^6$ samples dataset.
We then export the variation data from each subset to VCF
using \texttt{tskit vcf subset.ts | bgzip > subset.vcf.gz}
as the starting point for other tools.

We used bcftools version 1.18, Savvy 2.1.0, Genozip 5.0.26,
vcf2zarr 0.0.9, and Zarr-Python 2.17.2. 
All tools used default settings, 
unless otherwise stated.
All simulation-based benchmarks were performed on a 
dual CPU (Intel Xeon E5-2680 v2)
server with 256GiB of RAM running Debian GNU/Linux 11.
To ensure that the true effects of having data distributed over a large 
number of files were reported, benchmarks for Zarr and Savvy were 
performed on a cold disk cache by running 
\texttt{echo 3 | sudo tee /proc/sys/vm/drop\_caches} before each run.
The I/O subsystem used is based on a RAID 5 of 12 SATA hard drives.
For the CPU time benchmarks we measure the sum of the total user and
system times required to execute the full command (as reported by GNU
\texttt{time}) as well as elapsed wall-clock time. Total CPU
time is shown as a solid line, with wall-clock time as a dashed line 
of the same colour. In the case of pipelines, where some processing 
is conducted concurrently wall-clock time can be less than total
CPU (e.g.\ genozip in Fig~\ref{fig-whole-matrix-compute}). 
When I/0 costs are significant, wall-clock time can be greater 
than total CPU (e.g.\ Zarr and Savvy in Fig~\ref{fig-subset-matrix-compute}).
Each tool was instructed to use one thread, where the options
were provided. 
Where possible in pipelines we use uncompressed BCF
output (\texttt{-Ou}) to make processing 
more efficient~\citep{danecek2021twelve}.
We do not use BCF output in genozip because it is not supported
directly.

Because \texttt{bcftools +af-dist} requires the AF INFO field
and this is not kept in sync by \texttt{bcftools view}  
(although the AC and AN fields are), the subset calculation 
for Fig~\ref{fig-subset-matrix-compute} requires an additional step. 
The resulting pipeline is
\texttt{bcftools view -r REGION -S SAMPLESFILE -IOu BCFFILE | 
bcftools +fill-tags -Ou | bcftools +af-dist}. Genozip similarly
requires a \texttt{+fill-tags} step in the pipeline.

\section{Availability of source code and requirements}

The VCF Zarr specification is available on GitHub at 
\url{https://github.com/sgkit-dev/vcf-zarr-spec/}.
All source code for running benchmarks, analyses and creating 
"
plots in this article is available at
\url{https://github.com/sgkit-dev/vcf-zarr-publication}.
Vcf2zarr is freely available under the terms of the Apache 2.0 
license as part of the \texttt{bio2zarr} 
suite (\url{https://github.com/sgkit-dev/bio2zarr/}) 
and can be installed from the Python Package Index
(\url{https://pypi.org/project/bio2zarr/}).

\subsection{List of abbreviations}
% If abbreviations are used in the text they should be defined in the text at
% first use, and a list of abbreviations should be provided in alphabetical
% order.

\begin{itemize}
    \item ICF: Intermediate Columnar Format
    \item GWAS: Genome Wide Association Study
    \item PBWT: Positional Burrows-Wheeler Transform
    \item QC: Quality Control
    \item UKB: UK Biobank
    \item VCF: Variant Call Format
    \item WGS: Whole Genome Sequence
\end{itemize}

% \subsection{Competing Interests}
\subsection{Funding}
JK acknowledges support from the Robertson Foundation, 
NIH (research grants HG011395 and HG012473). 
JK and AM acknowledge the Bill \& Melinda Gates Foundation (INV-001927).

% \subsection{Author's Contributions}

\section{Acknowledgements}
This research was made possible through access to data in the National Genomic
Research Library, which is managed by Genomics England Limited (a wholly owned
company of the Department of Health and Social Care). The National Genomic
Research Library holds data provided by patients and collected by the NHS as
part of their care and data collected as part of their participation in
research. The National Genomic Research Library is funded by the National
Institute for Health Research and NHS England. The Wellcome Trust, Cancer
Research UK and the Medical Research Council have also funded research
infrastructure.

Computation used the Oxford Biomedical Research Computing (BMRC) facility, a
joint development between the Wellcome Centre for Human Genetics and the Big
Data Institute supported by Health Data Research UK and the NIHR Oxford
Biomedical Research Centre. The views expressed are those of the author(s) and
not necessarily those of the NHS, the NIHR or the Department of Health.

Genozip was used under the terms of the free Genozip Academic license.
Genozip was only used on simulated data, in compliance with
the ``No Commercial Data'' criterion.

\bibliography{paper}

\clearpage
\renewcommand\thefigure{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand\thetable{S\arabic{table}}
\setcounter{table}{0}

\section*{Supplementary Material}

\begin{figure}
\includegraphics{figures/whole-matrix-decode}
\caption{Genotype decoding performance.
Total CPU time required to decode genotypes into memory using the Zarr-Python
and Savvy C++ APIs for the data in Figure~\ref{fig-data-storage}.
This corresponds to a maximum rate of 1.2GiB/s for Zarr (Zstd + BitShuffle),
3.9 GiB/s Zarr (Zstd), and 6.6 GiB/s for Savvy. 
\label{fig-whole-matrix-decode}}
\end{figure}

\begin{figure}
\includegraphics{figures/subset-matrix-compute-supplemental}
\caption{Compute performance on a large subset of the genotype matrix.
Total CPU time required to run the af-dist calculation for
a subset of half of the samples and 10000 variants from the middle of the matrix
for the data in Figure~\ref{fig-data-storage}.
Genozip did not run for
$n > 10^4$ samples because it does not support a file to specify
sample IDs, and the command line was therefore too long for the shell
to execute. 
\label{fig-subset-matrix-compute-supplemental}}
\end{figure}
\begin{figure}

\includegraphics{figures/compression-compressor}
\caption{Effects of Blosc compression codec on compression ratio on call-level 
fields in 1000 Genomes data. 
In all cases compression level=7 was used, with a variant
chunk size of 10,000 and sample chunk size of 1,000.
Bit shuffle was used for call\_genotype, and no shuffle used for the other fields.
\label{fig-compression-compressor}}
\end{figure}

\begin{figure}
\includegraphics{figures/compression-shuffle}
\caption{Effects of Blosc shuffle settings on compression ratio on call-level 
fields in 1000 Genomes data. 
In all cases the zstd compressor with compression level=7 was used, with a variant
chunk size of 10,000 and sample chunk size of 1,000.
\label{fig-compression-shuffle}}
\end{figure}

\begin{figure}
\includegraphics{figures/compression-chunksize}
\caption{Effects of chunk sizes on compression ratio on call-level 
fields in 1000 Genomes data.
(A) Varying sample chunk size, holding variant chunk size fixed at 10,000.
(B) Varying variant chunk size, holding sample chunk size fixed at 1,000.
In all cases the zstd compressor with compression level=7 was used. Bit shuffle
was used for call\_genotype, and no shuffle used for the other fields.
Values are chosen to be evenly spaced on a linear scale 
between 100 and 2504 (the number of samples) in (A) and 
evenly spaced between 100 and 96514 on a log scale in (B). 
\label{fig-compression-chunksize}}
\end{figure}

\begin{figure}
\includegraphics{figures/compression-chunksize-finegrained.pdf}
\caption{Effects of sample chunk size on compression ratio on the call\_genotype
field in 1000 Genomes data.
The same analysis as in Fig~\ref{fig-compression-chunksize}, except we only
consider call\_genotype and we examine all sample chunk sizes from 
100 to 256. Distinct trend-lines emerge for odd, even and multiple-of-four
chunk sizes (shown by markers). The size of the final chunk also has a minor
effect (shown by colour).
\label{fig-compression-chunksize-finegrained}}
\end{figure}

\end{document}