### 2025-01-30 response-to-reviewers.tex
Add reviewer 1s comments

% LaTeX rebuttal letter example.
%
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
%
% Licensed under cc by-sa 3.0 with attribution required.

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
% define counters for reviewers and their points
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
\renewcommand{\thepoint}{\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule}
                  % \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Point~\thepoint} } ---\ }
   {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point}  \bigskip \noindent
    {\textbf{Reviewer~Point~\thepoint} } ---~#1\par }

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\  }
   {\medskip \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
    \ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
    \medskip \end{sf}}

\begin{document}

\section*{Response to the editor}
% General intro text goes here
Thank you for considering this manuscript for publication.  We have addressed
the points you and Reviewer 1 raised below in turn.

\reviewersection
\section*{Reviewer 1 comments}

\begin{point}
The authors present VCF Zarr, a specification that translates the variant call
format (VCF) data model into an array-based representation for the Zarr storage
format. They also present the `vcf2zarr` utility to convert large VCFs to Zarr.
They provide data compression and analysis benchmarks comparing VCF Zarr to
existing variant storage technologies using simulated genotype data. They also
present a case study on real world Genomics England aggV2 data.

The authors' benchmarks overall show that VCF Zarr has superior compression and
computational analysis performance at scale relative to data stored as
row-oriented VCF and that VCF Zarr is competitive with specialized storage
solutions that require similarly specialized tools and access libraries for
querying. An attractive feature is that VCF Zarr allows for variant annotation
workflows that do not require full dataset copy and conversion. Another key
point is that Zarr is a high-level spec and data model for the chunked storage
of n-d arrays, rather than a byte-level encoding designed specifically around
the genomic variant data type. I personally have used Zarr productively for
several applications unrelated to statistical genetics. While Zarr VCF mildly
underperforms some of the specialized formats (Savvy in compute, Genozip in
compression) in a few instances, I believe the accessibility, interoperability,
and reusability gains of Zarr make the small tradeoff well worthwhile.

Because Zarr has seen heavy adoption in other scientific communities like the
geospatial and Earth sciences, and is well integrated in the scientific Python
stack, I think it holds potential for greater reusability across the ecosystem.
As such, I think the VCF Zarr spec is a highly valuable if not overdue
contribution to an entrenched field that has recently been confronted by a
scalability wall.

Overall, the paper is clear, comprehensive, and well written. 
\end{point}
\begin{reply}
\end{reply}

\begin{point}
The benefits for large scientific datasets to be analysis-ready
cloud-optimized (ARCO) have been well articulated by Abernathey et al., 2021.
However, I do think that the "local"/HPC single-file use case is still
important and won't disappear any time soon, and for some file system use
cases, expansive and deep hierarchies can be performance limiting (this was
hinted at in one of the benchmarks). In this scenario would a large Zarr VCF
perform reasonably well (or even better on some file systems) via a single
local zip store?
\end{point}
\begin{reply}
\end{reply}

\begin{point}
The description of the intermediate columnar format (ICF) used by `vcf2zarr`
is missing some detail. At first I got the impression it might be based on
something like Parquet, but running the provided code showed that it consists
of a similar file-based chunk layout to Zarr. This should be clarified in the
manuscript.

\end{point}
\begin{reply}
\end{reply}

\begin{point}
The authors discuss the possibility of storing an index mapping genomic
coordinates to chunk indexes. Have Zarr-based formats in other fields like
geospatial introduced their own indexing approaches to take inspiration from?
\end{point}
\begin{reply}
\end{reply}

\begin{point}
Since VCF Zarr is still a draft proposal, it could be useful to indicate
where community discussions are happening and how potential new contributors
can get involved, if possible. This doesn't need to be in the paper per se, but
perhaps documented in the spec repo.
\end{point}
\begin{reply}
\end{reply}

\begin{point}
In the background: "For the representation to be FAIR, it must also be
accessible," -- A is for "accessible", so "also" doesn't make sense.
\end{point}
\begin{reply}
\end{reply}

\begin{point}
"There is currently no efficient, FAIR representation...". Just a nit and
feel free to ignore, but the solution you present is technically "current".
\end{point}
\begin{reply}
\end{reply}

\begin{point}
In Figure 2, the zarr line is occluded by the sav line and hard to see.
\end{point}
\begin{reply}
\end{reply}

\reviewersection
\section*{Reviewer 2 comments}

\begin{point}
Point
\end{point}
\begin{reply}
\end{reply}

\end{document}